---
title: "Monty Hall Problem"
author: "Tobias Holden"
date: "11/10/2021"
categories: ["R"]
tags: ["R Markdown", "puzzles"]
---

# Chorrrds

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)

theme_presentation<- function(base_size = 8, base_family = "") {
  # Starts with theme_grey and then modify some parts
  theme_bw(base_size = base_size, base_family = base_family) %+replace%
    theme(
      strip.background = element_blank(),
      strip.text.x = element_text(size = 12,colour="white"),
      strip.text.y = element_text(size = 12,colour="white"),
      axis.text.x = element_text(size=12,colour="white"),
      axis.text.y = element_text(size=12,colour="white",hjust=1),
      axis.ticks =  element_line(colour = "white"), 
      axis.title.x= element_text(size=12,colour="white"),
      axis.title.y= element_text(size=12,angle=90,colour="white"),
      axis.title.y.right= element_text(size=12,angle=90,colour="white"),
      panel.background = element_rect(fill="black"), 
      panel.border =element_blank(),  
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      panel.spacing = unit(1, "lines"), 
      plot.background = element_rect(fill="black"), 
      plot.title =element_text(size=12,colour="white"), 
      plot.margin = unit(c(0,1,0,1), "lines"),
      legend.background=element_rect(fill=NA, color=NA),
      legend.title=element_text(size=12,colour="white"),
      legend.text=element_text(size=12,colour="white"),
      legend.key = element_rect( fill = NA, colour=NA),
      legend.key.size = unit(c(1, 1), "lines"), 
      legend.position="bottom",
      legend.margin=margin(0,0,0,0),
      axis.line.x = element_line(color="white", size = 1),
        axis.line.y = element_line(color="white", size = 1)
    )
}
```

## Monty Hall Problem

1) Host draws a random # (1, 2, or 3) is  - this is the prize-winning number. The other 2 numbers are assigned "goat"
2) Contestant "chooses" a number 1, 2 or 3.
3) Host randomnly "reveals a goat", by by disclosing one of the non-winning numbers
4) Contestant either keeps their original choose (keep model) or switches (switch model) their number choice between the two remaining options.


```{r}
monty <- function(model = "keep", num = 100) {
  results <- c()
  mds <- c()
  ids <- c()
  for (md in model)
  {
    for (i in 1:num)
    {
      mds <- c(mds, md)
      ids <- c(ids, i)
      doors <- c("Door 1", "Door 2", "Door 3")
      car_door <- sample(doors, 1)
      goat_doors <- doors[doors != car_door]
      choice <- doors[runif(1, 2, 3)]
      if (md == "keep") {
        results <- c(results, choice == car_door)
      }
      if (md == "switch")
      {
        reveal <- sample(goat_doors[goat_doors != choice], 1)
        results <-
          c(results, doors[doors != choice & doors != reveal] == car_door)
      }
    }
  }
  
  #ids[ids>n] <- (ids[ids>n] %% (n+1) + 1)
  out <- data.frame(run = ids,
                    strategy = mds,
                    result = results)
  cumsum <- out %>% group_by(strategy) %>% summarise(z=cumsum(result))
  out$cum_win <- cumsum$z
  return(out)
  
}

montyDoors <- function(model = "keep", num = 100, d = 3, c=1) {
  results <- c()
  mds <- c()
  ids <- c()
  ncar <- c()
  ndoor <- c()
  for(nd in 3:d)
  {
    for(nc in 1:min(c,nd-2))
    {
  for (md in model)
  {
    for (i in 1:num)
    {
      ndoor <- c(ndoor,nd)
      ncar <- c(ncar,nc)
      mds <- c(mds, md)
      ids <- c(ids, i)
      doors <- paste("Door #", seq(1,nd))
      car_doors <- sample(doors, nc)
      goat_doors <- doors[!(doors %in% car_doors)]
      choice <- sample(doors,1)
      if (md == "keep") {
        results <- c(results, choice %in% car_doors)
      }
      if (md == "switch")
      {
        reveal <- sample(goat_doors[goat_doors != choice], 1)
        new_choice <- sample(doors[doors != choice & doors!=reveal],1)
        results <-c(results, new_choice %in% car_doors)
      }
    }
  }
    }
  }

  
  #ids[ids>n] <- (ids[ids>n] %% (n+1) + 1)
  out <- data.frame(run = ids,
                    n_car = ncar,
                    n_door = ndoor,
                    strategy = mds,
                    result = results)
  out <- out %>% group_by(strategy, n_door, n_car) %>% arrange(run) %>% summarise(run=run,result=result,cum_win=cumsum(result), final=NA)
  for (j in 1:nrow(out))
  {
    if(out$run[j]==num & out$strategy[j]=="switch")
    {
      d <- out$n_door[j]
      c <- out$n_car[j]
      hub <- subset(out, out$run==num & out$n_door==d & out$n_car ==c)
      out$final[j] <- round(hub$cum_win[hub$strategy=="switch"] / hub$cum_win[hub$strategy=="keep"],2)
    }
  }
  return(data.frame(out))
  
}





m <- montyDoors(model=c("keep","switch"), num = 1000, d=10 , c=8)
  
ggplot(m, aes(x=run)) +
  #geom_point(aes(y=as.numeric(result),color="Individual"),alpha=0.05) +
  #geom_jitter(aes(y=as.numeric(result),color="Individual"),alpha=0.1, height = 0.02) +
  #geom_segment(aes(xend=run, y=as.numeric(result),yend=cum_win/run, color=result), alpha=0.4,  size=0.2) +
  geom_path(aes(y=cum_win/run, color=strategy)) +
  facet_grid(n_door~n_car, labeller = label_both) +
  labs(color="", linetype="") +
  theme_presentation() +
  theme(legend.box="vertical", legend.direction = "horizontal", legend.justification = "left", legend.box.just = "left", legend.spacing = unit(-1,"lines"), axis.title.x=element_blank(), axis.text.x = element_blank()) +
  ylab("Winning") +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2))) +
  geom_text(data=m[!is.na(m$final),], aes(x=max(m$run)/2, y=0.8, label=final), x=max(m$run,na.rm=T)/ 2, y=0.8, color="yellow") 
  
```